---
title: "factorEx: Design and Analysis for Factorial Experiments"
output: 
  md_document:
    variant: markdown_github
---

# factorEx: Design and Analysis for Factorial Experiments

Authors: 

- [Naoki Egami](https://scholar.princeton.edu/negami/)
- [Brandon de la Cuesta](https://www.brandondelacuesta.com//)
- [Kosuke Imai](https://imai.fas.harvard.edu/)

References: 

 - de la Cuesta, Egami, and Imai. (2019+). [Improving the External Validity of Conjoint Analysis: The Essential Role of Profile Distribution.](https://scholar.princeton.edu/sites/default/files/negami/files/conjoint_profile.pdf)  (Working Paper)
 
  - Egami and Imai. (2019). [Causal Interaction in Factorial Experiments: Application to Conjoint Analysis.](https://scholar.princeton.edu/sites/default/files/negami/files/causalint.pdf) *Journal of the American Statistical Association*, Vol.114, No.526 (June), pp. 529–540.

## Installation Instructions
`factorEx` is available on CRAN and can be installed using:
```{r eval = FALSE}
install.packages("factorEx")
```

You can also install the most recent development version using the `devtools` package. First you have to install `devtools` using the following code. Note that you only have to do this once:
```{r eval=FALSE} 
if(!require(devtools)) install.packages("devtools")
```   
Then, load `devtools` and use the function `install_github()` to install `factorEx`:

```{r eval=FALSE}
library(devtools)
install_github("naoki-egami/factorEx", dependencies=TRUE)
```  

## Example 
`factorEx` implements the design-based and model-based approaches to estimate the pAMCE. 

### 1. Design-based Confirmatory Analysis
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
## Load the package and data
library(factorEx)
data("OnoBurden")
OnoBurden_data <- OnoBurden$OnoBurden_data
target_dist_marginal <- OnoBurden$target_dist_marginal

out_design_mar <- design_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security,
                           data = OnoBurden_data,
                           pair_id = OnoBurden_data$pair_id,
                           cluster_id = OnoBurden_data$id,
                           target_dist  = target_dist_marginal,
                           target_type = "marginal")
summary(out_design_mar, factor_name = "gender",  sample = TRUE)

target_dist_partial <- OnoBurden$target_dist_partial
partial_joint_name  <- list(c("gender", "age", "family"), "race", "party", c("experience", "pos_security"))
out_design_par <- design_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security,
                           data = OnoBurden_data,
                           pair_id = OnoBurden_data$pair_id,
                           cluster_id = OnoBurden_data$id,
                           target_dist  = target_dist_partial,
                           target_type = "partial_joint",
                           partial_joint_name = partial_joint_name)
summary(out_design_par, factor_name = "gender",  sample = TRUE)
```


```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
## Load the package and data
library(factorEx)
data("OnoBurden")
OnoBurden_data_pr <- OnoBurden$OnoBurden_data_pr
target_dist_marginal <- OnoBurden$target_dist_marginal

out_design_mar_pr <- design_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security, 
                                  factor_name  = c("gender"),
                                  data = OnoBurden_data_pr,
                                  pair_id = OnoBurden_data_pr$pair_id,
                                  cluster_id = OnoBurden_data_pr$id,
                                  target_dist  = target_dist_marginal,
                                  target_type = "marginal")
summary(out_design_mar_pr, factor_name = "gender")
summary(out_design_mar, factor_name = "gender")
```

