---
title: "factorEx: Design and Analysis for Factorial Experiments"
output: 
  md_document:
    variant: markdown_github
---

# factorEx: Design and Analysis for Factorial Experiments

**Description:** 

R package `factorEx` provides design-based and model-based estimators for the population average marginal component effects (the pAMCE) in factorial experiments, including conjoint analysis. The package also implements a series of recommendations offered in de la Cuesta, Egami, and Imai (2019+) and Egami and Imai (2019, JASA).

**Authors:** 

- [Naoki Egami](https://scholar.princeton.edu/negami/)
- [Brandon de la Cuesta](https://www.brandondelacuesta.com/)
- [Kosuke Imai](https://imai.fas.harvard.edu/)

**References:** 

 - de la Cuesta, Egami, and Imai. (2019+). [Improving the External Validity of Conjoint Analysis: The Essential Role of Profile Distribution.](https://scholar.princeton.edu/sites/default/files/negami/files/conjoint_profile.pdf)  (Working Paper)
 
  - Egami and Imai. (2019). [Causal Interaction in Factorial Experiments: Application to Conjoint Analysis.](https://scholar.princeton.edu/sites/default/files/negami/files/causalint.pdf) *Journal of the American Statistical Association*, Vol.114, No.526 (June), pp. 529–540.

## Installation Instructions
`factorEx` is available on CRAN and can be installed using:
```{r eval = FALSE}
install.packages("factorEx")
```

You can also install the most recent development version using the `devtools` package. First you have to install `devtools` using the following code. Note that you only have to do this once:
```{r eval=FALSE} 
if(!require(devtools)) install.packages("devtools")
```   
Then, load `devtools` and use the function `install_github()` to install `factorEx`:

```{r eval=FALSE}
library(devtools)
install_github("naoki-egami/factorEx", dependencies=TRUE)
```  

## Examples

- [Design-based Confirmatory Analysis](#design)
  - [Case 1: Use Marginal Distributions for Target Profile Distribution](#designMar)
  - [Case 2: Use Combination of Marginal and Partial Joint Distributions for Target Profile Distribution](#designJoint)

- [Model-based Exploratory Analysis](#model)

## (1) Design-based Confirmatory Analysis {#design}
Here, we use the conjoint experiment that randomized profiles according to the marginal population randomization design. 

### Case 1: Use Marginal Distributions for Target Profile Distributions {#designMar}
When using marginal distributions, `target_dist` should be a list and each element should have a factor name. Within each list, a `numeric` vector should have the same level names as those in `data`.
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
## Load the package and data
library(factorEx)
data("OnoBurden")

OnoBurden_data_pr <- OnoBurden$OnoBurden_data_pr # randomization based on marginal population design

# we focus on target profile distributions based on Democratic legislators. 
# See de la Cuesta, Egami, and Imai (2019+) for details.
target_dist_marginal <- OnoBurden$target_dist_marginal

target_dist_marginal
```

We can estimate the pAMCE with `design_pAMCE` with `target_type = "marginal"`. Use `factor_name` to specify for which factors we estimate the pAMCE. 
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
out_design_mar <- 
  design_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security,
               factor_name = c("gender", "age", "experience"),
               data = OnoBurden_data_pr,
               pair_id = OnoBurden_data_pr$pair_id,
               cluster_id = OnoBurden_data_pr$id,
               target_dist  = target_dist_marginal, target_type = "marginal")
summary(out_design_mar)
```

Use `plot` to visualize the estimated pAMCEs.
```{r eval = FALSE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
plot(out_design_mar, factor_name = c("gender", "experience"))
```

### Case 2: Use Combination of Marginal and Partial Joint Distributions for Target Profile Distribution {#designJoint}
The use of partial joint distributions is useful because it can relax the assumption of no three-way or higher-order interactions (see de la Cuesta, Egami, and Imai (2019+)).

When using a combination of marginal and partial joint distributions, `target_dist` should be a list and each element should be a `numeric` vector (if marginal) or an `array`/`table` (if partial joint). Then, use argument `partial_joint_name` to specify which factors are marginal and partial joints. In the following example, `c("gender", "age", "family")` has the partial joint distributions over the three factors. `race` and `party` are based on the marginal distributions, respectively. `c("experience", "pos_security")` has the partial joint distributions over the two factors. Within each list, a `numeric` vector or an `array`/`table` should have the same level names as those in `data`.

```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
target_dist_partial <- OnoBurden$target_dist_partial
target_dist_partial
partial_joint_name  <- list(c("gender", "age", "family"), "race", "party", c("experience", "pos_security"))
```

We can estimate the pAMCE with `design_pAMCE` with `target_type = "partial_joint"` and appropriate `partial_joint_name`. The function can use `factor_name` to specify for which factors we estimate the pAMCE. 
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
out_design_par <- 
      design_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security,
                   factor_name = c("gender", "age", "race"),
                   data = OnoBurden_data_pr,
                   pair_id = OnoBurden_data_pr$pair_id,
                   cluster_id = OnoBurden_data_pr$id,
                   target_dist  = target_dist_partial, target_type = "partial_joint",
                   partial_joint_name = partial_joint_name)
summary(out_design_par)
```

## (2) Model-based Exploratory Analysis {#model}
Here, we use the conjoint experiment that randomized profiles according to the uniform distribution and incorporate the target profile distribution in the analysis stage. 

```{r eval = TRUE, cache= TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
OnoBurden_data <- OnoBurden$OnoBurden_data # randomization based on uniform 

# due to large sample size, focus on "congressional candidates" for this example
OnoBurden_data_cong <- OnoBurden_data[OnoBurden_data$office == "Congress", ]

out_model <- 
      model_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security,
                   data = OnoBurden_data_cong, 
                  reg =  TRUE,
                   pair_id = OnoBurden_data_cong$pair_id,
                   cluster_id = OnoBurden_data_cong$id,
                   target_dist  = target_dist_marginal, target_type = "marginal")
summary(out_model, factor_name = c("gender"))
```

When `sample = TRUE`, the function also reports the AMCE based on the in-sample profile distributions (`sample AMCE`), which is the uniform AMCE in this example.
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
summary(out_model, factor_name = c("gender"), sample = TRUE)
```

Use `plot` to visualize the estimated pAMCEs. When `diagnose = TRUE`, it provides two diagnostic checks; specification tests and the check of bootstrap distributions. 
```{r eval = FALSE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
plot(out_model, factor_name = c("gender"), diagnose = TRUE)
```

In the model-based analysis, we can also decompose the difference between the pAMCE and the uniform AMCE. Use `effect_name` to specify which pAMCE we want to decompose. `effect_name` has two elements; the first is a factor name and  the second is a level name of interest. 
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
decompose_pAMCE(out_model, effect_name = c("gender", "Female"))
```

Or use `plot_decompose` to visualize the decomposition. 
```{r eval = FALSE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
plot_decompose(out_model, effect_name = c("gender", "Female"))
```
