---
title: "factorEx: Design and Analysis for Factorial Experiments"
output: 
  md_document:
    variant: markdown_github
---

# factorEx: Design and Analysis for Factorial Experiments

Description: 
R package `factorEx` provides design-based and model-based estimators for the population average marginal component effects (the pAMCE) in factorial experiments, including conjoint analysis. The package also implements a series of recommendations offered in de la Cuesta, Egami, and Imai (2019+).

Authors: 

- [Naoki Egami](https://scholar.princeton.edu/negami/)
- [Brandon de la Cuesta](https://www.brandondelacuesta.com//)
- [Kosuke Imai](https://imai.fas.harvard.edu/)

References: 

 - de la Cuesta, Egami, and Imai. (2019+). [Improving the External Validity of Conjoint Analysis: The Essential Role of Profile Distribution.](https://scholar.princeton.edu/sites/default/files/negami/files/conjoint_profile.pdf)  (Working Paper)
 
  - Egami and Imai. (2019). [Causal Interaction in Factorial Experiments: Application to Conjoint Analysis.](https://scholar.princeton.edu/sites/default/files/negami/files/causalint.pdf) *Journal of the American Statistical Association*, Vol.114, No.526 (June), pp. 529–540.

## Installation Instructions
`factorEx` is available on CRAN and can be installed using:
```{r eval = FALSE}
install.packages("factorEx")
```

You can also install the most recent development version using the `devtools` package. First you have to install `devtools` using the following code. Note that you only have to do this once:
```{r eval=FALSE} 
if(!require(devtools)) install.packages("devtools")
```   
Then, load `devtools` and use the function `install_github()` to install `factorEx`:

```{r eval=FALSE}
library(devtools)
install_github("naoki-egami/factorEx", dependencies=TRUE)
```  

## Example of Design-based Confirmatory Analysis
We use the conjoint experiment data based the marginal population randomization design. 

### Case 1. Target Profile Distributions as Combination of Marginal Distributions
When using marginal distributions, `target_dist` should be a list and each element should have a factor name. Within each list, a `numeric` vector should have the same level names as those in `data`.
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
## Load the package and data
library(factorEx)
data("OnoBurden")
OnoBurden_data_pr <- OnoBurden$OnoBurden_data_pr
target_dist_marginal <- OnoBurden$target_dist_marginal

target_dist_marginal
```

We can estimate the pAMCE with `design_pAMCE` with `target_type = "marginal"`. 
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
out_design_mar <- 
  design_pAMCE(formula = Y ~ gender + age + family + race + experience + pos_security,
               data = OnoBurden_data_pr,
               pair_id = OnoBurden_data_pr$pair_id,
               cluster_id = OnoBurden_data_pr$id,
               target_dist  = target_dist_marginal, target_type = "marginal")
summary(out_design_mar)
```

### Case 2. Target Profile Distributions as Combination of Marginal and Partial Joint Distributions
The use of partial joint distributions is useful because it can relax the assumption of no three-way or higher-order interactions (see de la Cuesta, Egami, and Imai (2019+)).

When using a combination of marginal and partial joint distributions, `target_dist` should be a list and each element should be a `numeric` vector (if marginal) or an `array`/`table` (if partial joint). Then, use argument `partial_joint_name` to specify which factors are marginal and partial joints. In the following example, `c("gender", "age", "family")` has the partial joint distributions over the three factors. `race` and `party` are based on the marginal distributions, respectively. `c("experience", "pos_security")` has the partial joint distributions over the two factors. Within each list, a `numeric` vector or an `array`/`table` should have the same level names as those in `data`.

```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
target_dist_partial <- OnoBurden$target_dist_partial
target_dist_partial
partial_joint_name  <- list(c("gender", "age", "family"), "race", "party", c("experience", "pos_security"))
```

We can estimate the pAMCE with `design_pAMCE` with `target_type = "partial_joint"` and appropriate `partial_joint_name`. The function can use `factor_name` to specify for which factors we estimate the pAMCE. 
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
out_design_par <- 
      design_pAMCE(formula = Y ~ gender + age + family + race + experience + party + pos_security,
                   factor_name = c("gender", "race"),
                   data = OnoBurden_data_pr,
                   pair_id = OnoBurden_data_pr$pair_id,
                   cluster_id = OnoBurden_data_pr$id,
                   target_dist  = target_dist_partial, target_type = "partial_joint",
                   partial_joint_name = partial_joint_name)
summary(out_design_par, factor_name = c("gender", "race"))
```
